### 2013-02-20 docs/parco13/ios-book-article.tex
Start directory for ParCo 2013 paper

\documentclass{IOS-Book-Article}

\usepackage{mathptmx}

%\usepackage{times}
%\normalfont
%\usepackage[T1]{fontenc}
%\usepackage[mtplusscr,mtbold]{mathtime}
%
\begin{document}
\begin{frontmatter}              % The preamble begins here.

%\pretitle{Pretitle}
\title{Instructions for the Preparation of an\\
Electronic Camera-Ready Manuscript in\\ \LaTeX}
\runningtitle{IOS Press Style Sample}
%\subtitle{Subtitle}

\author[A]{\fnms{Book Production} \snm{Manager}%
\thanks{Corresponding Author: Book Production Manager, IOS Press, Nieuwe Hemweg 6B,
1013 BG Amsterdam, The Netherlands; E-mail:
bookproduction@iospress.nl.}},
\author[B]{\fnms{Second} \snm{Author}}
and
\author[B]{\fnms{Third} \snm{Author}}

\runningauthor{B.P. Manager et al.}
\address[A]{Book Production Department, IOS Press, The Netherlands}
\address[B]{Short Affiliation of Second Author and Third Author}

\begin{abstract}
These instructions are designed for the Preparation of an Electronic
Camera-Ready Manuscript in \LaTeX{} and should be read carefully. If you
have any questions regarding the instructions, please contact the Book
Production Department, by e-mail: \textit{bookproduction@iospress.nl}
or fax: +31-20-6870039.
\end{abstract}

\begin{keyword}
electronic camera-ready manuscript\sep IOS Press\sep
\LaTeX\sep book\sep layout
\end{keyword}
\end{frontmatter}

\thispagestyle{empty}
\pagestyle{empty}

\section*{Introduction}
This document provides instructions for style and layout and how to submit the final
version. Although it was written for individual authors contributing to IOS Press books,
it can also be used by the author/editor preparing a monograph or an edited volume.

Authors should realize that the manuscript submitted by the volume editor to IOS
Press will be almost identical to the final, printed version that appears in the book,
except for the pagination and the insertion of running headlines. Proofreading as
regards technical content and English usage is the responsibility of the author.

A template file for \LaTeX2e is available from the Author's Corner on
www.iospress.nl. \LaTeX{} styles required for the \LaTeX{} template are also
available.\footnote{For authors using MS Word separate Instructions as well
as a Word template are available from the Author's Corner on
www.iospress.nl.}

\section{Typographical Style and Layout}

\subsection{Type Area}
The text output area is automatically set within an area 12.4 cm
horizontally and 20 cm vertically. Please do not use any
\LaTeX{} or \TeX{} commands that affect the layout or formatting of
your document (i.e. commands like \verb|\textheight|,
\verb|\textwidth|, etc.).

\subsection{Font}

The font type for running text (body text) is 10~point Times New Roman.
There is no need to code normal type (roman text). For literal text, please use
\texttt{type\-writer} (\verb|\texttt{}|)
or \textsf{sans serif} (\verb|\textsf{}|). \emph{Italic} (\verb|\emph{}|)
or \textbf{boldface} (\verb|\textbf{}|) should be used for emphasis.

\subsection{General Layout}
Use single (1.0) line spacing throughout the document. For the main
body of the paper use the commands of the standard \LaTeX{}
``article'' class. You can add packages or declare new \LaTeX{}
functions if and only if there is no conflict between your packages
and the \texttt{IOS-Book-Article}.

Always give a \verb|\label| where possible and use \verb|\ref| for cross-referencing.

\subsection{(Sub-)Section Headings}
Use the standard \LaTeX{} commands for headings: {\small \verb|\section|, \verb|\subsection|, \verb|\subsubsection|, \verb|\paragraph|}.
Headings will be automatically numbered.

Use initial capitals in the headings, except for articles (a, an, the), coordinate
conjunctions (and, or, nor), and prepositions, unless they appear at the beginning
of the heading.

\subsection{Footnotes and Endnotes}
Please keep footnotes to a minimum. If they take up more space than roughly 10\% of
the type area, list them as endnotes, before the References. Footnotes and endnotes
should both be numbered in arabic numerals and, in the case of endnotes, preceded by
the heading ``Endnotes''.

\subsection{References}

References to the literature should be mentioned in the main text by arabic numerals in
square brackets. Use the Citation-Sequence System, which means they are ``listed and
numbered in the sequence in which they are 1st cited. ($\ldots$) Subsequent citations of the
same document use the same numbers as that of its initial citation'' \cite{r1}.

As regards the content, form and punctuation of the References, if the volume
editor has not expressed a preference in this matter, authors should select the format
most appropriate to their article, and use it \textit{consistently}.

\section{Illustrations}

\subsection{General Remarks on Illustrations}
The text should include references to all illustrations. Refer to illustrations in the
text as Table~1, Table~2, Figure~1, Figure~2, etc., not with the section or chapter number
included, e.g. Table 3.2, Figure 4.3, etc. Do not use the words ``below'' or ``above''
referring to the tables, figures, etc.

Do not collect illustrations at the back of your article, but incorporate them in the
text. Position tables and figures at the top or bottom of a page, with at least 2 lines
extra space between tables or figures and the running text.

Illustrations should be centered on the page, except for small figures that can fit
side by side inside the type area. Tables and figures should not have text wrapped
alongside.

Place figure captions \textit{below} the figure, table captions \textit{above} the table.
Use bold for table/figure labels and numbers, e.g.: \textbf{Table 1.}, \textbf{Figure 2.},
and roman for the text of the caption. Keep table and figure captions justified. Center
short figure captions only.

The minimum \textit{font size} for characters in tables is 8 points, and for lettering in other
illustrations, 6 points.

On maps and other figures where a \textit{scale} is needed, use bar scales rather than
numerical ones of the type 1:10,000.

\subsection{Quality of Illustrations}
Use only Type I fonts for lettering in illustrations.

Include graphics in Encapsulated postscript (EPS) format. Do \textit{not} use illustrations
taken from the Internet. The resolution of images intended for viewing on a screen is
not sufficient for the printed version of the book.

If you are incorporating screen captures, keep in mind that the text may not be
legible after reproduction (using a screen capture tool, instead of the Print Screen
option of PC's, might help to improve the quality).

\subsection{Color Illustrations}
Please note, that illustrations will only be printed in color if the volume editor agrees to
pay the production costs for color printing. However, you should \textit{not} use color in
illustrations that must be printed in black and white, because this will greatly reduce the
print quality. (Note that in software the default often is color, so you may have to
change the settings for these illustrations.)

Illustrations that must be printed in colour should be enclosed as CMYK-encoded
EPS files.

\section{Equations}
Number equations consecutively, not section-wise. Place the numbers in parentheses at
the right-hand margin, level with the last line of the equation. Refer to equations in the
text as Eq. (1), Eqs. (3) and (5).

\section{Fine Tuning}

\subsection{Type Area}
\textbf{Check once more that all the text and illustrations are inside the type area and
that the type area is used to the maximum.} You may of course end a page with one
or more blank lines to avoid `widow' headings, or at the end of a chapter.

\subsection{Capitalization}
Use initial capitals in the title and headings, except for articles (a, an, the), coordinate
conjunctions (and, or, nor), and prepositions, unless they appear at the beginning of the
title or heading.

\subsection{Page Numbers and Running Headlines}
You do not need to include page numbers or running headlines. These elements will be
added by the publisher.

\section{Submitting the Manuscript}
Submit the following to the volume editor:

\begin{enumerate}
\item The main source file, and any other required files. Do not submit more than
one version of any item.

\item Identical high resolution PDF file with all fonts embedded. First produce a
Postscript file from \LaTeX{} with DVIPS version 5.56 or higher. The option
``-M'' (don't make fonts) should be indicated. Use Adobe Acrobat Distiller to
produce the PDF and choose the job option \textit{Press-Optimized}.
\end{enumerate}

\begin{thebibliography}{99}

\bibitem{r1}
\textit{Scientific Style and Format: The CBE manual for authors,
editors and publishers}. Style Manual Committee, Council of Biology Editors.
Sixth ed. Cambridge University Press, 1994.

\bibitem{r2}
L.U. Ante, Cem surgere: Surgite postquam sederitis, qui manducatis panem doloris,
\textit{Omnes} \textbf{13} (1916), 114--119.

\bibitem{r3}
T.X. Confortavit, \textit{Seras}, Portarum, New York, 1995.

\bibitem{r4}
P.A. Deus, Ater hoc et filius et mater praestet nobis,
\textit{Paterhoc} \textbf{66} (1993), 856--890.

\end{thebibliography}
\end{document}