### 2013-07-31 docs/parco13/conc.tex
Make figures fit, improvise conclusion

\section{Conclusion}

We have presented the integration of an MPI-based forest-of-octrees adaptive
meshing code, \pforest, with a numerical solver for hyperbolic
conservation laws, \clawpack/\manyclaw,
that implements threaded parallelism for a single compute patch
on the other.  All topological operations on the mesh are handled by \pforest,
while all numerical operations are handled by the \clawpack implementation.
This combination naturally enables a hybrid MPI/threaded parallelization.
We demonstrate scalability of this approach in various ways, where the MPI-only variant
scales weakly up to 16,384 codes, and strong scalability is linear for up to 256 cores
in the MPI-only and 128 in the threaded implementation